The fraction of polar aligned circumbinary disks

### 2025-04-14 src/tex/ms.tex
Eclipse timing variations for polar planets

We also discuss the limitations of our methods. First, we look at the effect of finite disk lifetime on the ability for a disk to complete its evolution to a stationary state. Disks with long evolutionary timescales remain with orbital parameters near their initial conditions for their entire lifetimes. We also look at limitations imposed by the fact that real disks are not geometrically narrow rings, but have some radial extent. Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too narrow of a ring. These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
\RR{
These results for $f_\text{polar}$ can be tested through population-level observations of circumbinary disk systems. In practice, however, the mutual inclination between the disk and its binary host can be difficult to constrain, and therefore polar disks can be difficult to distinguish from coplanar. Known misaligned disks rely on fortunate geometry to constrain the mutual inclination, such as an eclipse of a nearby binary in the case of V773 Tau B \citep{kenworthy2022}. However, $f_\text{polar}$ could be statistically constrained by looking for polar disks around young eclipsing binaries because the binary-observer inclination is known. Imaging of a face-on disk in such systems would confirm a polar configuration. However, for an edge-on disk it would be difficult to differentiate between a coplanar and polar due to a degeneracy in viewing angle. A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.
}

>>>>

We also discuss the limitations of our methods. First, we look at the effect of finite disk lifetime on the ability for a disk to complete its evolution to a stationary state. Disks with long evolutionary timescales remain with orbital parameters near their initial conditions for their entire lifetimes. We also look at limitations imposed by the fact that real disks are not geometrically narrow rings, but have some radial extent. Real disks are likely to break if their radial extent is too large, and will become gravitationally unstable if all of their mass in concentrated in too narrow of a ring. These two effects impose constraints on the relationship between $r_\text{in}$, $r_\text{out}$, and $j$.
\RR{
These results for $f_\text{polar}$ can be tested through population-level observations of circumbinary disk systems. In practice, however, the mutual inclination between the disk and its binary host can be difficult to constrain, and therefore polar disks can be difficult to distinguish from coplanar. Known misaligned disks rely on fortunate geometry to constrain the mutual inclination, such as an eclipse of a nearby binary in the case of V773 Tau B \citep{kenworthy2022}. However, $f_\text{polar}$ could be statistically constrained by looking for polar disks around young eclipsing binaries because the binary-observer inclination is known. Imaging of a face-on disk in such systems would confirm a polar configuration. However, for an edge-on disk it would be difficult to differentiate between a coplanar and polar due to a degeneracy in viewing angle. A carefully done survey of a large population of eclipsing binaries \citep[e.g., those found by the Kepler Mission][]{kirk2016} would likely constrain the fraction of polar circumbinary disks.

A similar method has been proposed by \citet{zhang2019} to distinguish polar from coplanar planets in eclipsing binary systems via eclipse timing variations. I contrast to the imaging method suggested previously, such a method would be less expensive, have fewer degeneracies, and have many more potential targets as it would not be restricted to young systems. Additionally, it would be most sensitive to planets with small separations from their host binaries, meaning the alignment timescale for the progenitor disk would be much shorter than for those that could be resolved via imaging. Together, these two methods could provide robust constraints of $f_\text{polar}$ across many regimes of host parameters and disk separation.
}
